---
title: "Machine Learning Group Project 1"
author: "Jesse DeBolt and Isaac Johnson"
date: "2023-01-25"
output: html_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(formatR)
library(moderndive)
library(caret)
library(class)
library(fastDummies)

wine <- read_rds("../pinot.rds")

```


From 1/24 class:
```{r}
# wino <- wine %>% 
#   rename_all(funs(tolower(.))) %>% 
#   rename_all(funs(str_replace_all(., "-", "_"))) %>% 
#   rename_all(funs(str_replace_all(., " ", "_"))) %>% 
#   mutate(old = str_detect(description, " old")) %>% 
#   mutate(village = str_detect(description, "[Vv]illage")) %>% 
#   mutate(northern = str_detect(description, "[Nn]orthern")) %>% 
#   select(-description) %>% 
#   drop_na(.)
# 
# head(wino) %>% 
#   select(1:8)
```

# 1/31 class:
## Create document term matrix
```{r}
library(tidytext)
data(stop_words)
head(stop_words, 25)$word


names(wine)[names(wine)=='id'] = 'ID'

df <- wine %>%
  unnest_tokens(word, description) %>% 
  anti_join(stop_words) %>% # get rid of stop words
  filter(word != "wine") %>%
  filter(word != "pinot") %>%
  count(ID, word) %>% #group by wine id and count
  group_by(ID) %>% 
  mutate(freq = n/sum(n)) %>% 
  mutate(exists = (n>0)) %>% 
  ungroup %>% 
  group_by(word) %>% 
  mutate(total = sum(n))

head(df, 10)

```

## Top words in database
```{r}
df %>% 
  count(word) %>%
  arrange(desc(n)) %>% 
  head(25)
```

## Pivot wide and rejoin with wine
```{r}
wino <- df %>% 
  filter(total > 1000) %>% #limit number of columns 
  pivot_wider(id_cols = ID, names_from = word, values_from = exists, values_fill = list(exists=0)) %>% 
  merge(select(wine,ID, province), all.y=TRUE) #%>% 

wino <- replace(wino, is.na(wino), FALSE)#if na, make FALSE

head(wino, 10) %>% 
  select(1:5,province)
```







```{r}
set.seed(504)
wine_index <- createDataPartition(wino$province, p = 0.8, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- knn(
  train = select(train,-province), 
  test = select(test,-province), 
  k=5, 
  cl = train$province, 
  prob = T)


```


## Confusion matrix

```{r tidy=T, strip.white=T, comment=""}
confusionMatrix(fit,factor(test$province))
```

## Basic model with parameter tuning

```{r}
control <- trainControl(method = "boot", number = 1)
fit <- train(province ~ .,
             data = train, 
             method = "knn",
             tuneLength = 15,
             trControl = control)
fit
```


## Confusion Matrix
```{r}
confusionMatrix(predict(fit, test),factor(test$province))
```


## With parameter tuning and subsampling

```{r}
fit <- train(province ~ .,
             data = train, 
             method = "knn",
             tuneLength = 15,
             metric = "Kappa",
             trControl = control)

fit
```

## Confusion Matrix
```{r}
confusionMatrix(predict(fit, test),factor(test$province))
```

## Tuning plot

```{r}
ggplot(fit, metric="Kappa")

```

## A new model
### Naive Bayes
```{r}
set.seed(504)
wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- train(province ~ .,
             data = train, 
             method = "naive_bayes",
             tuneGrid = expand.grid(usekernel = c(T,F), laplace = T, adjust = T),
                #can specify additional parameter/details for the algorithum to use
             metric = "Kappa",
             trControl = trainControl(method = "cv"))
fit
```

...now things are getting better.

## Confusion Matrix
```{r}
confusionMatrix(predict(fit, test),factor(test$province))
```
## Maybe we can find words associated with our sparse provinces?

```{r}
prov_words <- df %>% 
  left_join(select(wine, ID, province), by = "ID") %>% 
  count(province, word) %>%
  group_by(province) %>% 
  top_n(5,n) %>% 
  arrange(province, desc(n))

unique(prov_words$word)
```
