---
title: "Modeling Assignment 1"
Authors: Isaac Johnson & Jesse DeBolt
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Setup

```{r}
# add your libraries
library(tidyverse)
library(caret)
library(rpart)
# library(scales)
# library(janeaustenr)
# library(SnowballC)
#added
library(tidytext)
library(ggplot2)
library(scales)

wine = read_rds("../pinot.rds") 

```


## Feature Engineering
```{r}
# Feature Engineering based on top 50 words
wino <- wine %>% 
  rename_all(funs(tolower(.))) %>% 
  rename_all(funs(str_replace_all(., "-", "_"))) %>% 
  rename_all(funs(str_replace_all(., " ", "_"))) %>%
  mutate(fruit = str_detect(description, "fruit")) %>% 
  mutate(cherry = str_detect(description, "cherry")) %>%  
  mutate(flavors = str_detect(description, "[Ff]lavors")) %>% 
  mutate(black = str_detect(description, "black")) %>% 
  mutate(palate = str_detect(description, "palate")) %>% 
  mutate(red = str_detect(description, "red")) %>% 
  mutate(finish = str_detect(description, "finish")) %>% 
  mutate(tannis = str_detect(description, "tannis")) %>% 
  mutate(acidity = str_detect(description, "acidity")) %>%
  mutate(aromas = str_detect(description, "aromas")) %>%
  mutate(light = str_detect(description, "light")) %>%
  mutate(nose = str_detect(description, "nose")) %>% 
  mutate(drink = str_detect(description, "drink")) %>% 
  mutate(ripe = str_detect(description, "ripe")) %>%
  mutate(raspberry = str_detect(description, "raspberry")) %>%
  mutate(vineyard = str_detect(description, "vineyard")) %>%
  mutate(cranberry = str_detect(description, "cranberry")) %>%
  mutate(oak = str_detect(description, "oak")) %>%
  mutate(strawberry = str_detect(description, "strawberry")) %>% #25
  mutate(bodied = str_detect(description, "bodied")) %>%
  mutate(spice = str_detect(description, "spice")) %>%
  mutate(dark = str_detect(description, "dark")) %>%
  mutate(plum = str_detect(description, "plum")) %>%
  mutate(fruits = str_detect(description, "fruits")) %>% 
  mutate(texture = str_detect(description, "texture")) %>%
  mutate(tart = str_detect(description, "tart")) %>%
  mutate(bottling = str_detect(description, "bottling")) %>%
  mutate(fresh = str_detect(description, "fresh")) %>%
  mutate(rich = str_detect(description, "rich")) %>%
  mutate(soft = str_detect(description, "soft")) %>%
  mutate(earthy = str_detect(description, "earthy")) %>%
  mutate(berry = str_detect(description, "berry")) %>%
  mutate(notes = str_detect(description, "notes")) %>% 
  mutate(noir = str_detect(description, "noir")) %>%
  mutate(offers = str_detect(description, "offers")) %>%
  mutate(cola = str_detect(description, "cola")) %>%
  mutate(tea = str_detect(description, "tea")) %>%
  mutate(firm = str_detect(description, "firm")) %>%
  mutate(spicy = str_detect(description, "spicy")) %>%
  mutate(dried = str_detect(description, "dried")) %>%
  mutate(bright = str_detect(description, "bright")) %>%
  mutate(earth = str_detect(description, "earth")) %>%
  mutate(structure = str_detect(description, "structure")) %>%
  mutate(juicy = str_detect(description, "juicy")) %>% #50
  mutate(medium = str_detect(description, "medium")) %>%
  mutate(touch = str_detect(description, "touch")) %>%
  mutate(flavor = str_detect(description, "flavor")) %>%
  mutate(rose = str_detect(description, "rose")) %>%
  mutate(chocolate = str_detect(description, "chocolate")) %>%
  mutate(pomegranate = str_detect(description, "pomegranate")) %>%
  
  mutate(year = as.factor(year)) %>% 
  select(-c(description,id, year)) %>% #nb=.
#  select(-c(description,id)) %>% #knn=.5367
  drop_na(.)

```


## Specification
### First run based on top 50 words in wine dataset (kappa .7529)
```{r}
# specify the model to be used (i.e. KNN or Naive Bayes) and the tuning parameters used
start_time <- Sys.time()
# Code here
ctrl <- trainControl(method = "cv", number = 5)
set.seed(504) 

wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- train(province ~ .,
             data = train, 
             method = "rf", # Set method to random forest
             trControl = ctrl,
             tuneLength = 15,# Added tune length of 15
             metric = "Kappa")

confusionMatrix(predict(fit, test),factor(test$province))
fit

end_time <- Sys.time()
end_time - start_time

```


## Comparing words for NY versus California for standout NY words
```{r}
wtxt <- wine %>%
unnest_tokens(word, description) %>%
anti_join(stop_words) %>%
filter(str_detect(string = word, pattern = "[a-z+]")) %>% # get rid weird non alphas
filter(str_length(word)>3) %>% # get rid of strings shorter than 3 characters
group_by(word) %>%
mutate(total=n()) %>%
ungroup()
wtxt %>%
filter(province=="New_York" | province=="California") %>%
filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling"))) %>% 
filter(total < 100) %>%
group_by(province, word) %>%
count() %>%
group_by(province) %>%
mutate(proportion = n / sum(n)) %>%
pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>%
ggplot(aes(x = New_York, y = California, color = abs(New_York - California))) +
geom_abline(color = "gray40", lty = 2) +
geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
scale_x_log10(labels = percent_format()) +
scale_y_log10(labels = percent_format()) +
scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
theme(legend.position="none") +
labs(x = "New York", y = "California", title = "Words describing Pinot Noir from California and Oregon")

```


## Choosing 10 of the words in NY area of chart above
```{r}
# Words in NY not in Cali
# c(feathery, penetrating, island, cutting, loads, 
#                drying, astringent, lend, bramble, cassis, province)

```



## Feature Engineering II
### Adding NY words from comparision to California
```{r}
# Feature Engineering based on top 50 words
wino <- wine %>% 
  rename_all(funs(tolower(.))) %>% 
  rename_all(funs(str_replace_all(., "-", "_"))) %>% 
  rename_all(funs(str_replace_all(., " ", "_"))) %>%
  mutate(fruit = str_detect(description, "fruit")) %>% 
  mutate(cherry = str_detect(description, "cherry")) %>%  
  mutate(flavors = str_detect(description, "[Ff]lavors")) %>% 
  mutate(black = str_detect(description, "black")) %>% 
  mutate(palate = str_detect(description, "palate")) %>% 
  mutate(red = str_detect(description, "red")) %>% 
  mutate(finish = str_detect(description, "finish")) %>% 
  mutate(tannis = str_detect(description, "tannis")) %>% 
  mutate(acidity = str_detect(description, "acidity")) %>%
  mutate(aromas = str_detect(description, "aromas")) %>%
  mutate(light = str_detect(description, "light")) %>%
  mutate(nose = str_detect(description, "nose")) %>% 
  mutate(drink = str_detect(description, "drink")) %>% 
  mutate(ripe = str_detect(description, "ripe")) %>%
  mutate(raspberry = str_detect(description, "raspberry")) %>%
  mutate(vineyard = str_detect(description, "vineyard")) %>%
  mutate(cranberry = str_detect(description, "cranberry")) %>%
  mutate(oak = str_detect(description, "oak")) %>%
  mutate(strawberry = str_detect(description, "strawberry")) %>% #25
  mutate(bodied = str_detect(description, "bodied")) %>%
  mutate(spice = str_detect(description, "spice")) %>%
  mutate(dark = str_detect(description, "dark")) %>%
  mutate(plum = str_detect(description, "plum")) %>%
  mutate(fruits = str_detect(description, "fruits")) %>% 
  mutate(texture = str_detect(description, "texture")) %>%
  mutate(tart = str_detect(description, "tart")) %>%
  mutate(bottling = str_detect(description, "bottling")) %>%
  mutate(fresh = str_detect(description, "fresh")) %>%
  mutate(rich = str_detect(description, "rich")) %>%
  mutate(soft = str_detect(description, "soft")) %>%
  mutate(earthy = str_detect(description, "earthy")) %>%
  mutate(berry = str_detect(description, "berry")) %>%
  mutate(notes = str_detect(description, "notes")) %>% 
  mutate(noir = str_detect(description, "noir")) %>%
  mutate(offers = str_detect(description, "offers")) %>%
  mutate(cola = str_detect(description, "cola")) %>%
  mutate(tea = str_detect(description, "tea")) %>%
  mutate(firm = str_detect(description, "firm")) %>%
  mutate(spicy = str_detect(description, "spicy")) %>%
  mutate(dried = str_detect(description, "dried")) %>%
  mutate(bright = str_detect(description, "bright")) %>%
  mutate(earth = str_detect(description, "earth")) %>%
  mutate(structure = str_detect(description, "structure")) %>%
  mutate(juicy = str_detect(description, "juicy")) %>% #50
  mutate(medium = str_detect(description, "medium")) %>%
  mutate(touch = str_detect(description, "touch")) %>%
  mutate(flavor = str_detect(description, "flavor")) %>%
  mutate(rose = str_detect(description, "rose")) %>%
  mutate(chocolate = str_detect(description, "chocolate")) %>%
  mutate(pomegranate = str_detect(description, "pomegranate")) %>%
  # NY not in California words  
  mutate(feathery = str_detect(description, "feathery")) %>%
  mutate(penetrating = str_detect(description, "penetrating")) %>%
  mutate(island = str_detect(description, "island")) %>%
  mutate(cutting = str_detect(description, "cutting")) %>%
  mutate(loads = str_detect(description, "loads")) %>%
  mutate(drying = str_detect(description, "drying")) %>%
  mutate(astringent = str_detect(description, "astringent")) %>%
  mutate(lend = str_detect(description, "lend")) %>%
  mutate(bramble = str_detect(description, "bramble")) %>%
  mutate(cassis = str_detect(description, "cassis")) %>%
  # NY not in California words 
  mutate(year = as.factor(year)) %>% 
  select(-c(description,id, year)) %>% #nb=.
#  select(-c(description,id)) %>% #knn=.5367
  drop_na(.)

```


## Specification
### Second run based on adding ten words in NY but not Cali (kappa .7644)
```{r}
# specify the model to be used (i.e. KNN or Naive Bayes) and the tuning parameters used

start_time <- Sys.time()
# Code here
ctrl <- trainControl(method = "cv", number = 5)
set.seed(504) 

wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- train(province ~ .,
             data = train, 
             method = "rf", # Set method to random forest
             trControl = ctrl,
             tuneLength = 15,# Added tune length 
             metric = "Kappa")

confusionMatrix(predict(fit, test),factor(test$province))
fit

end_time <- Sys.time()
end_time - start_time

```


# Final model testing
## Best model
```{r}
# Here are a few lines to inspect your best model. Add some comments about optimal hyperparameters.
print(fit)
print(fit$bestTune)

```


## Re-fit and evaluation
```{r}
# the "method" below should match the one you chose above. 
start_time <- Sys.time()
# Code here
set.seed(1504) # I will choose a different seed for evaluation

wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

# example spec for knn
fit_final <- train(province ~ .,
             data = train, 
             method = "rf", # Set method to random forest
             tuneLength = 15,# Added tune length 
             tuneGrid=fit$bestTune) 
# The last line means we will fit a model using the best tune parameters your CV found above.

confusionMatrix(predict(fit_final, test),factor(test$province))


end_time <- Sys.time()
end_time - start_time

```







## experiment 1 weighting more CA more OR kappa .764
```{r}
#summary count of each province
table(wino$province)

# Create a vector of weights for each province
province_weights <- ifelse(train$province %in% c("California", "Oregon"), 2, 1)

# Pass the weights vector to the `train()` function
fit <- train(province ~ .,
             data = train, 
             method = "rf", 
             trControl = ctrl,
             tuneLength = 15, 
             metric = "Kappa",
             weights = province_weights)


```

```{r}

confusionMatrix(predict(fit, test),factor(test$province))
fit
```

## experiment 2 weighting more CV more Marl, more Burg (kappa .763) 14 min
```{r}

#timer
start_time <- Sys.time()
# Create a vector of weights for each province
province_weights <- ifelse(train$province %in% c("Burgundy", "Marlborough, Casablanca_Valley"), 5, 1)

# Pass the weights vector to the `train()` function
fit <- train(province ~ .,
             data = train, 
             method = "rf", 
             trControl = ctrl,
             tuneLength = 15, 
             metric = "Kappa",
             weights = province_weights)

#timer end
end_time <- Sys.time()
end_time - start_time
```

```{r}

confusionMatrix(predict(fit, test),factor(test$province))
fit
```
##visualize importance of variables
```{r}

featurePlot(x = train[, 1:3],
            y = as.factor(train$province),
            plot = "box")
            


featurePlot(x = train[, 1:3],
            y = as.factor(train$province),
            plot = "box",
            strip=strip.custom(par.strip.text = list(cex = .7)),
            scales = list(x = list(relation = "free"),
                          y = list(relation = "free")))
```
## Experiment #3 weighted fruit=ture (kappa .7571) 14 min
```{r}
#timer
start_time <- Sys.time()
# Create a vector of weights for each province
province_weights <- ifelse(train$fruit %in% c("TRUE"), 5, 1)

# Pass the weights vector to the `train()` function
fit <- train(province ~ .,
             data = train, 
             method = "rf", 
             trControl = ctrl,
             tuneLength = 15, 
             metric = "Kappa",
             weights = province_weights)

#timer end
end_time <- Sys.time()
end_time - start_time

confusionMatrix(predict(fit, test),factor(test$province))
fit
 
```



## Experiment #4 weighted fruit cherry flavors black palate (kappa .7516) 14 min
```{r}
#timer
start_time <- Sys.time()
# Create a vector of weights for each province
province_weights <- ifelse(train$fruit %in% c("TRUE"), 5, 1) |
    ifelse(train$cherry %in% c("TRUE"), 5, 1) |
    ifelse(train$flavors %in% c("TRUE"), 5, 1) |
    ifelse(train$black %in% c("TRUE"), 5, 1) |
    ifelse(train$palate %in% c("TRUE"), 5, 1)

# Pass the weights vector to the `train()` function
fit <- train(province ~ .,
             data = train, 
             method = "rf", 
             trControl = ctrl,
             tuneLength = 15, 
             metric = "Kappa",
             weights = province_weights)

#timer end
end_time <- Sys.time()
end_time - start_time

confusionMatrix(predict(fit, test),factor(test$province))
fit

#fruit cherry Fflavors black palate red finish. tannins acidity aromas light
 
```



## Experiment #5 weighted price>90 (kappa .756) 14 min
```{r}
#timer
start_time <- Sys.time()
# Create a vector of weights for each province
province_weights <- ifelse(train$price>90, 5, 1)

# Pass the weights vector to the `train()` function
fit <- train(province ~ .,
             data = train, 
             method = "rf", 
             trControl = ctrl,
             tuneLength = 15, 
             metric = "Kappa",
             weights = province_weights)

#timer end
end_time <- Sys.time()
end_time - start_time

confusionMatrix(predict(fit, test),factor(test$province))
fit

#fruit cherry Fflavors black palate red finish. tannins acidity aromas light
 
```


```{r}

ggplot(wine, aes(x = (price))) +
    geom_histogram(bins=10) +
    xlim(0, 100)

ggplot(wine, aes(x = (points))) +
    geom_histogram(bins=10)


```

